Information Closure Theory of Consciousness

### 2020-04-10 ICT_Frontiers_rev202001.tex
Local update
Updates from Overleaf

@@

    %     \begin{definition}
    %     Given a stochastic process $X$ called the universe process, a \emph{C-process} is a coarse-graining $Y$ of $X$ such that the following two conditions are satisfied (see Fig.~\ref{fig:fullgraph}):
    %     \begin{enumerate}
    %     \item $Y$ is informationally closed to $X$
    %     \item there exists a pair $(S,E)$ of coarse-grainings of $X$ called the \textit{system} $S$ and its \textit{environment} $E$  such that 
    %     \begin{itemize}
    %         \item $Y$ is a coarse-graining of $S$,
    %         \item the state space $\mathcal{X}$ of $X$ is equal to the Cartesian product of the state spaces $\mathcal{S}$ and $\mathcal{E}$ of processes $S$ and $E$ respectively, formally $\mathcal{X}=\mathcal{S}\times\mathcal{E}$, and 
    %         \item $E$ is the environment process of $S$ with respect to which $Y$ is maximally NTIC:
    %         \begin{align}
    %         f_E:=\argmax_{\{\E||\E|\leq |\X|\},g:\X \rightarrow \E} NTIC(g(X)\rightarrow Y) \end{align}
    %     %     $Y$ is NTIC to $E$, formally:
    %     % \begin{align}
    %     %     NTIC_t(E\rightarrow Y) >0
    %     % \end{align}
    %     \end{itemize} 
    %     \end{enumerate}
    %   \end{definition}

@@

        \begin{hypothesis}
        A process $Y$ is conscious if and only if it is a C-process of some process $X$. Also the content of consciousness $C_t^{Content}$ at time $t$ is the state $y_t$ of the C-process at time $t$ and the level of consciousness $C_t^{Level}$ is the degree of NTIC of the process to the environment i.e. $NTIC_t(E\rightarrow Y)$:
        \begin{align}
            C_t^{Content} &= y_t \label{eq:cContent}\\
            C_t^{Level} &= NTIC_t(E\rightarrow Y) \label{eq:cLevel}
        \end{align}
        \end{hypothesis}

@@

    	    %% not monotonic
    		\linelabel{line:non-monotonic}It is important to note that NTIC can be a non-monotonic function of the scale of coarse-graining
    % 		\footnote{We have known that NTIC is not a non-monotonic function of the scale of coarse-graining. However, we currently have limited understanding about this relationship. Future work is needed and may involve relevant informational techniques (e.g.,partial information decomposition).}. 
    		At the finest scale we consider the whole universe $X$ as the process $Y$. then, since $Y$ is a coarse graining of $S$ we have $Y=S=X$.  
    % 		the universe should be informationally closed. 
    		In this case the environment $E$ corresponding to the universe seen as a system is the constant coarse-graining\footnote{Recall that, for a system with state space $\S$ the environment state space $\E$ must be such that $\X=\S \times \E$. I $\S =\X$ then we need $\E$ with $\X \times \E=\X$ such that $\E$ must be a singleton set. All coarse-grainings mapping $\X$ to a singleton set are constant over $\X$.} and therefore the mutual information $I(E_t;Y_{t+1})=0$ and the transfer entropy $I(Y_{t+1};E_t|Y_t)$ are zero. Then NTIC of the universe with respect to its environment is zero and $X$ can never be a C-process.
    % 		For simplicity let us assume that we choose $Y=S$ in the following as well. This lets

    		If we now increase the scale of $Y$, this allows $S$ to also reduce in scale and therefore $E$ can become more and more fine-grained. This means that the mutual information $I(E_t;Y_{t+1})$ between $E$ and $Y$ can at least potentially become positive. Up to the point where $E$ accounts for half of the bits of $X$ and $S$ for the other half (and we have $Y=S$) the maximally possible mutual information $I(E_t;Y_{t+1})$ increases. Even further refining $E$ then again leads to a reduction of the maximally possible $I(E_t;Y_{t+1})$.

            % ==================================== %
            %    WHY DO WE NEED the FUCKING $Y$    %
            % ==================================== %
            % 2020/04/07 01:57
    		% Why does Definition 2 have a fucking $Y$? It seems the whole thing could be defined wihtout referring to the underlying $S$ and directly use $Y$'s own environment (together with which it corresponds to $X$). We would always have a refinement of $Y$ that could be called $S$ to gether with its environment. But this is alwyas the case so it doesn't restrict the definition of $Y$ at all. (Especially if we can just choose $S=Y$). The only thing that makes the current definition of $Y$ more general is that it may be that $Y$ is only closed with respect to a coarsening of its own environment and not to its own environment itself. This is a waekening of the restriction compared to requiring that $Y$ be NTIC with respect to its own environment and IC with respect to $X$. Actually no, since $Y$ is IC with respect $X$ it is IC with respect to everything including to any coarsening of its own environment. Since the mutual information between $Y$ and its environment is greater or equal to the mutual information between $Y$ and a coarsening of its environment $Y$ must also be NTIC with respect to its own environment. 
    		% ==================================== %

@@

            % ---------------------------------------------------------------------------- %
            %                              Dynamical Boundary                              %
            % ---------------------------------------------------------------------------- %
            % \subsection{Dynamical Boundary}
            % Our theory predicts that the boundary of a conscious system is dynamical. The boundary is determined by the NTIC process. This implies that when the structure of interaction between variables changes the physical boundaries supporting consciousness should also change. If some coarse-grained variables encode environmental information, they may become necessary components to form high NTIC. In such case, these coarse-grained variables are "recruited" inside the boundary of conscious processes. Note that, due to synergistic information, even recruiting some new variables may largely increase the level of NTIC. 
            % This may explain why consciousness is tightly associated with binding and integration of information in human perception and cognition. Another crucial implication is that the same neural substrates may not be always involved in conscious processing rather it depends on whether it creates high NTIC with other members in the neural system. \needfig{Maybe I need a figure here as well}

>>>>

			\subsection{Informational Closure and Causality}
			A surprising result of the definition of information flow $Jt(E\rightarrow Y)$ (Eq. \ref{eq:InformationFlow}) is that it does not involve causal dependency between $E$ and $Y$.
			Consider the case already noted by \citet{BERTSCHINGER.2006} where the sensory process $\hat{e_t}$ is deterministic (or close to it) and the system just copies the sensory states. Then the system can potentially achieve positive values of NTIC and informational closure with respect to the environment. 
			This scenario is called "passive adaptation" \citet{BERTSCHINGER.2006}. Counterintuitively, in such case, the future state is totally determined by the environment however, the information flow $Jt(E\rightarrow Y)$ is (or nearly) zero.

			@ However, in more realistic cases, the environment is often partial obseriable. The system cannot copy the entire environment state because it observes only part of it.

            We show in Appendix \ref{sec:appendix} that whenever the environment has itself more predictable dynamics than the observations there is potential for a predictive process to achieve higher NTIC than a copying system or any system that only processes its last observations without taking into account other internal memory. In this case is called "modelling" in \cite{BERTSCHINGER.2006}

            In such case, the system can potentially achieve higher NTIC than simply copy the sensory value.

            We can see both senatios are critical in ICT.

            % Similar to \citet{BERTSCHINGER.2006} we will refer to systems of the latter type as \textit{passive}. We will call all other systems that achieve positive NTIC \textit{predictive}. Note that the predictive process might need to be more complex than the system that only copies its observations.

@@

    % ==================================== %
    % martin.biehl: 
    % doesn't work because the information closrue with respect to $X$ implies full closrure with respect to every 
    % other process. This means all conscious processes are fully closed. This contradicts what we say below.
    % Apr 7, 2020 3:23 AM
    % ==================================== %
    %     \begin{definition}
    %     Given a stochastic process $X$ called the universe process, a \emph{C-process} is a coarse-graining $Y$ of $X$ such that the following two conditions are satisfied (see Fig.~\ref{fig:fullgraph}):
    %     \begin{enumerate}
    %     \item $Y$ is informationally closed to $X$
    %     \item there exists a pair $(S,E)$ of coarse-grainings of $X$ called the \textit{system} $S$ and its \textit{environment} $E$  such that 
    %     \begin{itemize}
    %         \item $Y$ is a coarse-graining of $S$,
    %         \item the state space $\mathcal{X}$ of $X$ is equal to the Cartesian product of the state spaces $\mathcal{S}$ and $\mathcal{E}$ of processes $S$ and $E$ respectively, formally $\mathcal{X}=\mathcal{S}\times\mathcal{E}$, and 
    %         \item $E$ is the environment process of $S$ with respect to which $Y$ is maximally NTIC:
    %         \begin{align}
    %         f_E:=\argmax_{\{\E||\E|\leq |\X|\},g:\X \rightarrow \E} NTIC(g(X)\rightarrow Y) \end{align}
    %     %     $Y$ is NTIC to $E$, formally:
    %     % \begin{align}
    %     %     NTIC_t(E\rightarrow Y) >0
    %     % \end{align}
    %     \end{itemize} 
    %     \end{enumerate}
    %   \end{definition}
    % ====================================%

@@

@@

    	    %% not monotonic
    		\linelabel{line:non-monotonic}It is important to note that NTIC can be a non-monotonic function of the scale of coarse-graining. At the finest scale we consider the whole universe $X$ as the process $Y$. then, since $Y$ is a coarse graining of $S$ we have $Y=S=X$. In this case the environment $E$ corresponding to the universe seen as a system is the constant coarse-graining\footnote{Recall that, for a system with state space $\S$ the environment state space $\E$ must be such that $\X=\S \times \E$. I $\S =\X$ then we need $\E$ with $\X \times \E=\X$ such that $\E$ must be a singleton set. All coarse-grainings mapping $\X$ to a singleton set are constant over $\X$.} and therefore the mutual information $I(E_t;Y_{t+1})=0$ and the transfer entropy $I(Y_{t+1};E_t|Y_t)$ are zero. Then NTIC of the universe with respect to its environment is zero and $X$ can never be a C-process.

    		If we now increase the scale of $Y$, this allows $S$ to also reduce in scale and therefore $E$ can become more and more fine-grained. This means that the mutual information $I(E_t;Y_{t+1})$ between $E$ and $Y$ can at least potentially become positive. Up to the point where $E$ accounts for half of the bits of $X$ and $S$ for the other half (and we have $Y=S$) the maximally possible mutual information $I(E_t;Y_{t+1})$ increases. Even further refining $E$ then again leads to a reduction of the maximally possible $I(E_t;Y_{t+1})$.
            % ==================================== %
            %    WHY DO WE NEED the FUCKING $Y$    %
            % ==================================== %
            % 2020/04/07 01:57
    		% Why does Definition 2 have a fucking $Y$? It seems the whole thing could be defined wihtout referring to the underlying $S$ and directly use $Y$'s own environment (together with which it corresponds to $X$). We would always have a refinement of $Y$ that could be called $S$ to gether with its environment. But this is alwyas the case so it doesn't restrict the definition of $Y$ at all. (Especially if we can just choose $S=Y$). The only thing that makes the current definition of $Y$ more general is that it may be that $Y$ is only closed with respect to a coarsening of its own environment and not to its own environment itself. This is a waekening of the restriction compared to requiring that $Y$ be NTIC with respect to its own environment and IC with respect to $X$. Actually no, since $Y$ is IC with respect $X$ it is IC with respect to everything including to any coarsening of its own environment. Since the mutual information between $Y$ and its environment is greater or equal to the mutual information between $Y$ and a coarsening of its environment $Y$ must also be NTIC with respect to its own environment. 
    		% ==================================== %

@@
